Counting infinitely by oritatami co-transcriptional folding

### 2019-08-28 paper.tex
w

\title{Counting infinitely by Oritatami co-transcriptional folding}

%\titlerunning{}
\author{
Kohei Maruyama\inst{1}\thanks{Primary corresponding author (\email{k.maruyama@uec.ac.jp})} \and
Shinnosuke Seki\inst{1,2}\thanks{Secondary corresponding author (\email{s.seki@uec.ac.jp})}
}
\institute{
The University of Electro-Communications, 
%Graduate School of Informatics and Engineering, 
1-5-1 Chofugaoka, Chofu, Tokyo, 1828585, Japan \and
\'{E}cole Normale Superi\'{e}ure de Lyon, 46 all\'{e}e d'Italie, 69007, Lyon, France
}

@@

An \textit{oritatami system} (OS) $\Xi$ is a tuple $(\Sigma, R, \delta, \alpha, \sigma, w)$, where $\Sigma$ and $R$ are defined as above, while
\begin{itemize}
\item a positive integer $\delta$ called the \textit{delay}, 
\item a positive integer $\alpha$ called the \textit{arity}, %←arityの行だけ要らない？
\item an initial $R$-valid conformation $\sigma \in C_{\le \alpha}(\Sigma)$ called the \textit{seed},
\item a (possibly infinite) \textit{transcript} $w \in \Sigma^* \cup \Sigma^\omega$, which is to be folded upon the seed by stabilizing beads of $w$ one at a time so as to minimize energy collaboratively with the succeeding $\delta{-}1$ nascent beads. 
\end{itemize}
%
The energy of a conformation $C = (P, w, H)$, denoted by $\Delta G(C)$, is defined to be ${-}|H|$; the more bonds a conformation has, the more stable it gets. 
The set $\mathcal{F}(\Xi)$ of conformations \textit{foldable} by the system $\Xi$ is recursively defined as: the seed $\sigma$ is in $\mathcal{F}(\Xi)$; and provided that an elongation $C_i$ of $\sigma$ by the prefix $w[1..i]$ be foldable (i.e., $C_0 = \sigma$), its further elongation $C_{i+1}$ by the next bead $w[i+1]$ is foldable if 
\begin{equation}\label{eq:OS_CF}
C_{i+1} \in \argmin_{
\substack{
C \in \mathcal{C}_{\le \alpha} s.t. \\
C_i \xrightarrow{R}_{w[i+1]}C \\
}
}
\min \Bigl\{ \Delta G(C') \Bigm|
C \xrightarrow{R}^*_{w[i+2...i+k]}C', k\le \delta, C' \in \mathcal{C}_{\le \alpha}
\Bigr\}.
\end{equation}
%
Then we say that the bead $w[i+1]$ and the bonds it forms are \textit{stabilized} according to $C_{i+1}$. 
The easiest way to understand this stabilization process should be the video available at \href{https://www.dailymotion.com/video/x3cdj35}{\tt https://www.dailymotion.com/video/x3cdj35}, in which the Turing universal oritatami system by Geary et al. \cite{GeMeScSe2018}, whose delay is 3, is running. 
Note that an arity-$\alpha$ oritatami system cannot fold any conformation of arity larger than $\alpha$. %←
A conformation foldable by $\Xi$ is \textit{terminal} if none of its elongations is foldable by $\Xi$. 
%The oritatami system $\Xi$ is \textit{deterministic} if for all $i \ge 0$, there exists at most one $C_{i+1}$ that satisfies \eqref{eq:OS_CF}. 
%A deterministic oritatami system folds into a unique terminal conformation. 
%An oritatami system with the empty rule set just folds into an arbitrary elongation of its seed nondeterministically. 
%Thus, the rule set is reasonably assumed non-empty.

\begin{figure}[tb]
\begin{tabular}{c}
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.6\linewidth]{fig/svg/Fnt_1.pdf}
 \end{minipage}

 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.6\linewidth]{fig/svg/Fnb_1.pdf}

 \end{minipage}
 \end{tabular}
\ \\
\ \\
\begin{tabular}{c}
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.6\linewidth]{fig/svg/F0_1.pdf}
 \end{minipage}

 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.6\linewidth]{fig/svg/F1_1.pdf}

 \end{minipage}
 \end{tabular}
 \caption{All the 4 bricks of module F: The 2 bricks at the top are for zigs while the 2 bottom are for zags. }
 \label{fig:formatters}
\end{figure}

\subsection{General idea}
Between two consecutive overflows, the proposed system behaves in the same way as the finite binary counter of Geary et al. \cite{GeMeScSe2019}.
Its transcript folds in a zigzag manner macroscopically (downward in figures throughout this paper).
A zig, folding from right to left, increments the current value of the counter by 1.
The succeeding zag, folding from left to right, formats the incremented value for the sake of next zig and copies it downward.
Unlike the counter of Geary et al., when a zig encounters an overflow, it does not abort but extends the current value by 1 bit.

The transcript of our counter is periodic.
Its period 1-2-3- $\cdots$ -132 is semantically divided into the following 4 subsequences, called \textit{modules}:
\begin{itemize}
\item 1--30: Format module or F
\item 31--66: Left-Turn module or L
\item 67--96: Half-Adder module or H
\item 97--132: Right-Turn module  or R
\end{itemize}
Modules are to play their roles in expected environments by folding into respective conformations which should be pairwise-distinct enough to be distinguishable by other modules transcribed later.
Such expected conformations are called a \textit{brick}.
For example, module F encounters only the four environments shown in Fig.~\ref{fig:formatters} where it takes the four bricks, respectively.

\begin{figure}[tb]
\begin{tabular}{c}
 \begin{minipage}{0.3\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/H00_1.pdf}
 \end{minipage}

 \begin{minipage}{0.3\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/H01_1.pdf}

  \begin{minipage}{0.3\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/He1_1.pdf}
 \end{minipage}
 \end{tabular}

\begin{tabular}{c}
\ \\
 \begin{minipage}{0.3\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/H10_1.pdf}
 \end{minipage}

 \begin{minipage}{0.3\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/H11_1.pdf}
 \end{minipage}

  \begin{minipage}{0.3\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Hn_1.pdf}

\begin{figure}[tb]
\begin{tabular}{c}
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.8\linewidth]{fig/svg/Lt_3.pdf}
 \end{minipage}

 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.8\linewidth]{fig/svg/Lbc_3.pdf}

@@

\begin{tabular}{c}
\ \\
\ \\
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.8\linewidth]{fig/svg/Ltre_3.pdf}
 \end{minipage}

  \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Lbe_3.pdf}
   \vspace*{18mm}

@@

@@

\paragraph{Seed and Encoding.}
The initial counter value is encoded as $b_{k-1}b_{k-2} \cdots b_1b_0$ in binary on the seed in the following format
\begin{equation} \label{eq:zagencoding}
64{-}65{-}66{-}\left( \prod^0_{i = k-1} \bigl(  w_{Hn} w_{Rb} w_{Fb_i} w_{Lbn} \bigr) \right) w_{Hn}
\end{equation}
where $w_{Hn}, w_{Rb}, w_{Fb_i}, w_{Lbn}$ are sequences of bead types exposed downward by modules H, R, F, L when they fold into bricks Hn, Rb, F$b_i$, Lbn, respectively, which can be found in Figs.~\ref{fig:halfadders}, \ref{fig:rightturns}, \ref{fig:formatters}, and \ref{fig:leftturns}.
The seed for $k = 1$ and $b_0 = 0$ is colored in purple in Fig.~\ref{fig:counter1stzig}.
%$w_{Hn} = 67{-}76{-}77{-}78{-}79{-}88{-}89{-}90{-}91{-}96$ \\
%$w_{Rb} = 97{-}102{-}103{-}108{-}109{-}114{-}115{-}120{-}121{-}126{-}127{-}132$\\
%$w_{Lbn} = 31{-}36{-}37{-}42{-}43{-}48{-}49{-}54{-}55{-}64{-}65{-}66$

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{fig/svg/CounterEx5_1.pdf}
\caption{
The first zig pass.
The seed is encoded with "0" same as bottom of F0(
\textcircled{\scriptsize 1}{-}\textcircled{\scriptsize 10}{-}\textcircled{\scriptsize 11}{-}\textcircled{\scriptsize 12}{-}\textcircled{\scriptsize 13}{-}\textcircled{\scriptsize 22}{-}\textcircled{\scriptsize 23}{-}\textcircled{\scriptsize 24}{-}\textcircled{\scriptsize 25}{-}\textcircled{\scriptsize 30}),
 and the counter increments it by giving a carry.
Module H outputs "1" as a sum and cancels the carry.
}
\label{fig:counter1stzig}
\end{figure}

>>>>

\title{Counting infinitely by oritatami co-transcriptional folding}

%\titlerunning{}
\author{
Kohei Maruyama\inst{1}\and
Shinnosuke Seki\inst{1,2}
}
\institute{
The University of Electro-Communications, 
%Graduate School of Informatics and Engineering, 
1-5-1 Chofugaoka, Chofu, Tokyo, 1828585, Japan \email{\{k.maruyama, s.seki\}@uec.ac.jp}\and
\'{E}cole Normale Superi\'{e}ure de Lyon, 46 all\'{e}e d'Italie, 69007, Lyon, France
}

@@

An \textit{oritatami system} (OS) $\Xi$ is a tuple $(\Sigma, R, \delta, \alpha, \sigma, w)$, where $\Sigma$ and $R$ are defined as above, while
\begin{itemize}
\item a positive integer $\delta$ called \textit{delay}, 
\item a positive integer $\alpha$ called \textit{arity}, %←arityの行だけ要らない？
\item an initial $R$-valid conformation $\sigma \in C_{\le \alpha}(\Sigma)$ called the \textit{seed},
\item a (possibly infinite) \textit{transcript} $w \in \Sigma^* \cup \Sigma^\omega$, which is to be folded upon the seed by stabilizing beads of $w$ one at a time so as to minimize energy collaboratively with the succeeding $\delta{-}1$ nascent beads. 
\end{itemize}
%
The energy of a conformation $C = (P, w, H)$, denoted by $\Delta G(C)$, is defined to be ${-}|H|$; the more bonds a conformation has, the more stable it gets. 
The set $\mathcal{F}(\Xi)$ of conformations \textit{foldable} by the system $\Xi$ is recursively defined as: the seed $\sigma$ is in $\mathcal{F}(\Xi)$; and provided that an elongation $C_i$ of $\sigma$ by the prefix $w[1..i]$ be foldable (i.e., $C_0 = \sigma$), its further elongation $C_{i+1}$ by the next bead $w[i+1]$ is foldable if 
\begin{equation}\label{eq:OS_CF}
C_{i+1} \in \argmin_{
\substack{
C \in \mathcal{C}_{\le \alpha} s.t. \\
C_i \xrightarrow{R}_{w[i+1]}C \\
}
}
\min \Bigl\{ \Delta G(C') \Bigm|
C \xrightarrow{R}^*_{w[i+2...i+k]}C', k\le \delta, C' \in \mathcal{C}_{\le \alpha}
\Bigr\}.
\end{equation}
%
Then we say that the bead $w[i+1]$ and the bonds it forms are \textit{stabilized} according to $C_{i+1}$. 
The easiest way to understand this stabilization process should be the video available at \href{https://www.dailymotion.com/video/x3cdj35}{\tt https://www.dailymotion.com/video/x3cdj35}, in which the Turing universal oritatami system by Geary et al. \cite{GeMeScSe2018}, whose delay is 3, is running. 
Note that an arity-$\alpha$ oritatami system cannot fold any conformation of arity larger than $\alpha$. %←
A conformation foldable by $\Xi$ is \textit{terminal} if none of its elongations is foldable by $\Xi$. 
%The oritatami system $\Xi$ is \textit{deterministic} if for all $i \ge 0$, there exists at most one $C_{i+1}$ that satisfies \eqref{eq:OS_CF}. 
%A deterministic oritatami system folds into a unique terminal conformation. 
%An oritatami system with the empty rule set just folds into an arbitrary elongation of its seed nondeterministically. 
%Thus, the rule set is reasonably assumed non-empty.

\begin{figure}[tb]
\begin{tabular}{c}
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.8\linewidth]{fig/svg/Fnt_1.pdf}
 \end{minipage}

 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.8\linewidth]{fig/svg/Fnb_1.pdf}

 \end{minipage}
 \end{tabular}
\ \\
\ \\
\begin{tabular}{c}
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.8\linewidth]{fig/svg/F0_1.pdf}
 \end{minipage}

 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.8\linewidth]{fig/svg/F1_1.pdf}

 \end{minipage}
 \end{tabular}
 \caption{All the four bricks of module F: The two bricks at the top, \texttt{Fnt} and \texttt{Fnb}, are for zigs while the others, \texttt{F0} and \texttt{F1}, are for zags. }
 \label{fig:formatters}
\end{figure}

\begin{figure}[tb]
\begin{tabular}{c}
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Lt_3.pdf}
 \end{minipage}

 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Lbc_3.pdf}

@@

\begin{tabular}{c}
\ \\
\ \\
 \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Ltre_3.pdf}
 \end{minipage}

  \begin{minipage}{0.5\linewidth}
  \centering
   \includegraphics[width=\linewidth]{fig/svg/Lbe_3.pdf}
   \vspace*{18mm}

@@

\subsection{General idea}
Between two consecutive overflows, the proposed system behaves in the same way as the finite binary counter proposed by Geary et al. \cite{GeMeScSe2019}.
Its transcript folds in a zigzag manner macroscopically (downward in figures throughout this paper).
A zig, folding from right to left, increments the current value of the counter by 1.
The succeeding zag, folding from left to right, formats the incremented value for the sake of next zig and copies it downward.
Unlike the existing counter, when a zig encounters an overflow, it does not abort but rather extends the current value by 1 bit.

The transcript of our counter is periodic.
Its period 1-2-3- $\cdots$ -132 is semantically divided into the following 4 subsequences, called \textit{modules}:
\begin{itemize}
\item 1--30: Format module or F
\item 31--66: Left-Turn module or L
\item 67--96: Half-Adder module or H
\item 97--132: Right-Turn module  or R
\end{itemize}
The transcript can be hence represented as $(FLHR)^*$ at the module level. Modules are to play their roles in expected environments by folding into respective conformations which should be pairwise-distinct enough to be distinguishable by other modules transcribed later.
Such expected conformations are called a \textit{brick}.
For example, module F encounters only the four environments shown in Fig.~\ref{fig:formatters} where it takes the four bricks, respectively.

\begin{figure}[tb]
\begin{tabular}{c}
 \begin{minipage}{0.33\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/H00_1.pdf}
 \end{minipage}

 \begin{minipage}{0.33\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/H01_1.pdf}

  \begin{minipage}{0.33\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/He1_1.pdf}
 \end{minipage}
 \end{tabular}

\begin{tabular}{c}
\ \\
 \begin{minipage}{0.33\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/H10_1.pdf}
 \end{minipage}

 \begin{minipage}{0.33\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/H11_1.pdf}
 \end{minipage}

  \begin{minipage}{0.33\linewidth}
  \centering
   \includegraphics[width=0.9\linewidth]{fig/svg/Hn_1.pdf}

@@

\paragraph{Seed and Encoding.}
The initial counter value is encoded as $b_{k-1}b_{k-2} \cdots b_1b_0$ in binary on the seed in the following format
\begin{equation} \label{eq:zagencoding}
64{-}65{-}66{-}\left( \prod^0_{i = k-1} \bigl(  w_{Hn} w_{Rb} w_{Fb_i} w_{Lbn} \bigr) \right) w_{Hn}
\end{equation}
where $w_{Hn}= 67{-}76{-}77{-}78{-}79{-}88{-}89{-}90{-}91{-}96$, $w_{Rb}=97{-}102{-}103{-}108{-}$ $109{-}114{-}115{-}120{-}121{-}126{-}127{-}132$, $w_{F0} = 1{-}10{-}11{-}12{-}13{-}22{-}23{-}24{-}$ $25{-}30$, $w_{F1}=1{-}22{-}23{-}24{-}25{-}26{-}27{-}28{-}29{-}30$, $w_{Lbn}= 31{-}36{-}37{-}42{-}$ $43{-}48{-}49{-}54{-}55{-}64{-}65{-}66$ are sequences of bead types exposed downward by modules H, R, F, L when they fold into bricks \texttt{Hn}, \texttt{Rb}, \texttt{F}$b_i$, \texttt{Lbn}, respectively, which can be found in Figs.~\ref{fig:halfadders}, \ref{fig:rightturns}, \ref{fig:formatters}, and \ref{fig:leftturns}.
For instance, the seed for $k = 1$ and $b_0 = 0$ is colored in purple in Fig.~\ref{fig:counter1stzig}.
%$w_{Hn} = 67{-}76{-}77{-}78{-}79{-}88{-}89{-}90{-}91{-}96$ \\
%$w_{Rb} = 97{-}102{-}103{-}108{-}109{-}114{-}115{-}120{-}121{-}126{-}127{-}132$\\
%$w_{Lbn} = 31{-}36{-}37{-}42{-}43{-}48{-}49{-}54{-}55{-}64{-}65{-}66$

\subsection{Brick level overview}
Starting from the seed, this system cyclically transits four phases: zig ($\leftarrow$), left carriage-return ($\hookrightarrow$), zag ($\rightarrow$), and right carriage-return ($\hookleftarrow$). 
The prefix $(FLHR)^k F$ of the transcript folds into the first zig (recall that $k$ is the bit-width of the initial value). 
In zigs in general, all the instances of modules F and H fold into bricks of width 10 and height 3, while all the instances of  L and R fold into bricks of width 12 and height 3. 
Zigs thus turn out to be a linear structure of height 3. 
In the process of the prefix folding into the first zig, its $i$-th instance of $H$ is transcribed right below $b_{i-1}$ encoded on the seed in the format (2) so that the H can ``read" $b_{i-1}$. 
After the whole prefix thus has folded into the first zig, the next L is transcribed right below Turn Signal, which lets the L fold into a special brick for left carriage-return if the zig ended at the top (this occurs when $b_{k-1} b_{k-2} \cdots b_0 < 1^k$) (see Fig.~\ref{fig:counter1stzag}). 
We should note that this special brick \texttt{Lcre} is provided with another Turn Signal for the sake of next left carriage-return.
Having been thus carriage-returned, the succeeding subsequence $(HRFL)^k H$ of the transcript folds into the first zag. 
Even in zags instances of F and H fold into bricks of width 10 and height 3, while those of L and R fold into bricks of width 12 and height 3. 
As a result, zags turn out to be a linear structure of height 3. 
More importantly, instances of H and F are aligned vertically and alternately into columns (see Figs.~\ref{fig:counter1stzig}-\ref{fig:overflowex1}), $i$-th of which from the right propagates the ($i{-}1$)-th bit of the counter value downward. 
After the whole subsequence has folded into the first zag, an instance of $R$ is transcribed and folded into a special brick \texttt{Rcr} for carriage-return due to the turn signal 125-124-123-122, which occurs also at the bottom of \texttt{Rcr} for the sake of next right carriage-return. 
This amounts to one cycle of the phase transition.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{fig/svg/CounterEx5_1.pdf}
\caption{
The first zig.
0 is encoded as an initial value below the seed in the format \eqref{eq:zagencoding} with $k = 1$ (1-bit width).
Being fed with carry, the zig increments the value.
Module H outputs 1, or more precisely a sequence of bead types which shall be interpreted as 1 in the next zag and reformatted, as a sum and cancels the carry.
}
%\textcircled{\scriptsize 1}{-}\textcircled{\scriptsize 10}{-}\textcircled{\scriptsize 11}{-}\textcircled{\scriptsize 12}{-}\textcircled{\scriptsize 13}{-}\textcircled{\scriptsize 22}{-}\textcircled{\scriptsize 23}{-}\textcircled{\scriptsize 24}{-}\textcircled{\scriptsize 25}{-}\textcircled{\scriptsize 30},

\label{fig:counter1stzig}
\end{figure}
